---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "fig/README-"
)
```

# fillgaze

The goal of fillgaze is to provide helper functions for interpolating missing
eyetracking data.

## Installation

You can install fillgaze from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("tjmahr/fillgaze")
```

## Package overview

This package was created in response to a very strange file of eyetracking data.

```{r example, message = FALSE}
df <- readr::read_csv("inst/test-gaze.csv")
```

Here is the problem with this file:

```{r, fig.height = 3}
library(dplyr, warn.conflicts = FALSE)
library(ggplot2)

ggplot(head(df, 40)) + 
  aes(x = Time - min(Time)) + 
  geom_hline(yintercept = 0, size = 2, color = "white") + 
  geom_point(aes(y = GazeX, color = "GazeX")) +
  geom_point(aes(y = GazeY, color = "GazeY")) + 
  labs(x = "Time (ms)", y = "Screen location (pixels)", 
       color = "Variable")
```

Every second or third point is incorrectly placed offscreen, indicated by a
negative pixel values for the gaze locations. It is physiologically impossible 
for a person's gaze to oscillate so quickly and with such magnitude (the gaze 
is tracked on a large screen display).

We would like to interpolate spans of missing data using 
neighboring points. That's the point of this package.
The steps to solve the problem involve:

- [x] Converting offscreen values into proper `NA` values.
- [x] Identifying and describing gaps of missing values (streaks of 
  successive `NA`s).
- [x] Interpolating the values in a gap.
- [ ] Writing tests to confirm that everything works as expected. :innocent:

### Setting values in several columns to `NA`

We need to mark offscreen points as properly missing data.
`set_values_to_na()` takes a dataframe and named filtering predicates. 
Here's the basic usage.

```{r, eval = FALSE}
set_values_to_na(dataframe, {col_name} = {function to determine NA values})
```

The values that return `TRUE` for each function are replaced with `NA` values.
For example, `set_values_to_na(df, var1 = ~ .x < 0)` would:

* look for the column `var1` in the dataframe, 
* check which values of `.x < 0` are true where `.x` is a placeholder/pronoun 
  for the values in `df$var1`,
* and replace those values where the test is `TRUE` with `NA`.

```{r}
library(fillgaze)
original_df <- df

df <- df %>% 
  set_values_to_na(
    GazeX = ~ .x < -100, 
    GazeY = ~ .x < -100, 
    LEyeCoordX = ~ .x < -.1, 
    LEyeCoordY = ~ .x < -.1,
    REyeCoordX = ~ .x < -.1, 
    REyeCoordY = ~ .x < -.1)

# Before and after on some of the GazeX values
data_frame(before = head(original_df$GazeX), after = head(df$GazeX))
```

Now, those offscreen points will not be plotted because they are `NA`.

```{r, fig.height = 3}
last_plot() %+% head(df, 40)
```

### Finding gaps in the data

We can use `find_gaze_gaps()` to locate the gaps in a column of data. This
function mostly is used internally. Users are not expectedly to routinely use
this function, but I cover it here because the function for filling gaps relies
on the data in this dataframe.

```{r, width = 120}
find_gaze_gaps(df, GazeX) %>% 
  print(width = 120)
```

Each row describes a gap in the column. 

* `start_row` and `end_row` contain row numbers of nearest non-`NA` values. 
* `na_rows` is the number of successive `NA`s in the gap.
* `start_value`and `end_value` contain the nearest non-`NA` values. 
* `change_value` is the difference between `start_value` and `end_value`.

The function also measure the duration of the gap (`change_time`). By default,
it uses row numbers (`.rowid`) to measure duration. We can use an explicit
column to use as the measure of time.

```{r, width = 120}
find_gaze_gaps(df, GazeX, time_var = Time) %>% 
  print(width = 120)
```

The function also respects dplyr grouping, so that e.g., false gaps are not found
between trials.

```{r}
df %>% 
  group_by(Trial) %>% 
  find_gaze_gaps(GazeX)
```

```{r}
## todo: unit tests
```



### Interpolating values in gaps

`fill_gaze_gaps()` will fill in the gaps in selected columns. We can set limits 
on which gaps are filled:

- `max_na_rows`: don't fill gaps with more than successive `NA` rows than `max_na_rows`
- `max_duration`: don't fill gaps with a duration larger than `max_duration` 
- `max_sd`: don't fill gaps where the relative change in the variable is more than `max_sd` standard deviations in magnitude

```{r}
df <- df %>% 
  group_by(Trial) %>% 
  fill_gaze_gaps(GazeX, time_var = Time, max_na_rows = 5)
```

```{r}
## todo: unit tests
```


In this example, only `GazeX` has been interpolated. The median value is used.
We can compare the results with the `GazeY` column.

```{r}
df %>% select(Time:GazeY)
```

`fill_gaze_gaps()` also works with the variable selection helpers from dplyr/tidyselect.

```{r}
df <- df %>% 
  fill_gaze_gaps(GazeX, GazeY, matches("EyeCoord"), 
                 time_var = Time, max_na_rows = 5, max_sd = 2) %>% 
  ungroup()
df
```

```{r, fig.height = 3}
was_offscreen <- (original_df$GazeX < -100) %>% 
  ifelse("Interpolated", "Original\nRaw Data")

last_plot() %+% 
  head(df, 40) + 
  aes(alpha = head(was_offscreen, 40), shape = head(was_offscreen, 40)) + 
  scale_alpha_discrete(name = "Point", range = c(.3, 1)) + 
  labs(shape = "Point")
```



